---
title: "FF_datacheck"
date: "13 September 2016"
output: html_document
---
```{r, echo=FALSE, message = FALSE, warning = FALSE}
setwd("~/Documents/frequent flyer/data_flyers")
library(survival)
library(xtable)
library(Hmisc)
library(gridBase, gridExtra)
library(lattice)
library(lubridate)
library(dplyr)
```

Ned passed these two files to Franck for NLP analysis:
```{r}
#Loading notes
dcsum <- read.csv("dischargeSummariesClean.csv")
#nursum <- read.csv("nursingNotesClean.csv")
```

We want to find out the level of agreement within each pair of annotators/operators; ie between:
1. ETM and JTW (Ned and Joy)
2. JF and JW (John Foote and Jonathan Welt)

```{r}
#Discharge summaries annotated by each operator:expect ETM & JTW and JF & JW to have annotated the same number of notes - but big differences. ?why
summary(dcsum$operator)

#Number of discharge annotations:
nrow(dcsum)
#Number of unique hospital admissions annotated:
length(unique(dcsum$Hospital.Admission.ID))
#dcsum$Hospital.Admission.ID is complete
summary(is.na(dcsum$Hospital.Admission.ID))
#dcsum$ICU.ID is complete
summary(is.na(dcsum$ICU.ID))
length(unique(dcsum$ICU.ID))
#dcsum$Subject.ID is complete
summary(is.na(dcsum$Subject.ID))
length(unique(dcsum$Subject.ID))
#dcsum$Chart.time is complete
summary(is.na(dcsum$Chart.time))
length(unique(dcsum$Chart.time))

#Subset notes by operator
ETM <- dcsum[dcsum$operator == "ETM",]
JTW <- dcsum[dcsum$operator == "JTW",]
JF <- dcsum[dcsum$operator == "JF",]
JW <- dcsum[dcsum$operator == "JW",]

#There are duplicated notes annotated by each operator:
ETMdup <- ETM[duplicated(ETM$ICU.ID),]; nrow(ETMdup)
JTWdup <- JTW[duplicated(JTW$ICU.ID),]; nrow(JTWdup)
JFdup <- JF[duplicated(JF$ICU.ID),]; nrow(JFdup)
JWdup <- JW[duplicated(JW$ICU.ID),]; nrow(JWdup)
#Create subset with all the duplicates from dcsum
dcsum_dup <- dcsum[duplicated(dcsum$ICU.ID),]; nrow(dcsum_dup)
length(unique(dcsum$ICU.ID))
```

#Questions for Ned:

1. Did you just rbind all the results file together to get the combined discharge and nursing .csv files?
Ned - I did rbind it-- I loaded a dataframe, bound it, loaded a new one, bound it, etc.

2. How come Jonathan annotated 560 notes and John Foote only 338 notes?
Joy - We are missing some notes from everyone. John is looking through his old computer to see if he could dig up the files. 

3. Where are the notes Patrick annotated? I recall he annotated some new notes and reviewed some of our unsure notes. Are the new notes Patrick annotated listed under JW, hence he appeared to have annotated 200+ more notes than John? I don't see Patrick as an operator.
Ned doesn't know where they are so Patrick will have to send the annotated unsured notes to me again.
Also after doing some digging around - every annotator (ETM, JTW, JF, JW) is supposed to have annotated 650 notes (including duplicates sent out by Ned).

#Interim Summary:

```{r, echo=FALSE}
print(paste("Summary: It would appear that Ned only sent us notes that have all of ICU.ID, Hospital.Admission.ID, Subject.ID and Chart.time to be annotated. We have a total", nrow(dcsum), "discharge annotations from all annotators, including any duplicated annotations by each operator. However, we are still missing some annotated notes from each operator - there should be 2600 annotated discharge notes if we have all the results files back. With what we have so far, there are", length(unique(dcsum$ICU.ID)), "unique discharge summary notes annotated, from", length(unique(dcsum$Hospital.Admission.ID)), "unique hospital or icu admissions. There are only", length(unique(dcsum$Subject.ID)), "unique patients. Each operator annotated some discharge notes (ETM =", nrow(ETMdup), ", JTW =", nrow(JTWdup), ", JF =", nrow(JFdup), ", JW =", nrow(JWdup), ") more than once. This duplication consists of a total of", nrow(dcsum_dup), "of the annotated notes. Could the NLP people use these intra-operator duplicated notes to get a sense intra-operator variability?", sep = " "))
```

# Remaining missing notes and notes with only one annotator

```{r}
#setwd("~/Documents/frequent flyer/data_flyers")

#read in hospital admission ids of notes annotated by only one person from AllDischargeFinal.csv
#"SingleAnnotatorNotes.csv" was pulled by Yeran
annot1 <- unique(read.csv("SingleAnnotatorNotes.csv")$Hospital.Admission.ID)
length(annot1)

#read in hadm.id of remaining missing notes batches
miss1 <- unique(read.csv("ejDis16Mar16.csv")$hadm.id); length(miss1)
miss2 <- unique(read.csv("JJdis22OCT15.csv")$hadm.id); length(miss2)
miss3 <- unique(read.csv("jjDis16Mar16.csv")$hadm.id); length(miss3)
remainmiss <- unique(c(miss1, miss2, miss3)); length(remainmiss)

#Remain missing ID as a subset of single annotator ID
missINannot1 <- remainmiss %in% annot1
missID <- remainmiss[missINannot1]
length(missID)

miss1_id <- miss1[miss1 %in% annot1]; length(miss1_id)
miss2_id <- miss2[miss2 %in% annot1]; length(miss2_id) #strange that all notes in "JJdis22OCT15.csv" seem to have been annotated twice already but they shouldn't have been...
miss3_id <- miss3[miss3 %in% annot1]; length(miss3_id)

#single annotator ID not in Remain missing ID
annot1INmiss <- !(annot1 %in% remainmiss)
missID2 <- annot1[annot1INmiss];length(missID2)
#Where/which batch did these 43 IDs come from then?

#I'll try subsetting out notes annotated once from AllDischargeFinal27Sep16.csv myself.
alldis <- read.csv("AllDischargeFinal27Sep16.csv")
by_hadmid <- group_by(alldis, Hospital.Admission.ID)
annot_num <- data.frame(summarise(by_hadmid, numNote <- n()))
str(annot_num)
colnames(annot_num) <- c("hadm.id", "num.notes")
annot_once <- filter(annot_num, num.notes == 1); str(annot_once)
annot1.id <- unique(annot_once$hadm.id); length(annot1.id)


#Check which note still needs annotating 
Miss1_id <- miss1[miss1 %in% annot1.id]; length(Miss1_id)
# So Jonathan can just keep annotating his missing batch
Miss2_id <- miss2[miss2 %in% annot1.id]; length(Miss2_id) 
# So all the notes in this batch has been annotated at least twice already
# but how...?
Miss3_id <- miss3[miss3 %in% annot1.id]; length(Miss3_id)
#Will split this batch up to two as Patrick and David requested.
Miss4_id <- annot1.id[!(annot1.id %in% remainmiss)]; length(Miss4_id)
# So where do these 41 missing notes come from?

#See which notes are in miss2
length(unique(miss2))
batch2 <- alldis[alldis$Hospital.Admission.ID %in% miss2,]
nrow(batch2)
length(unique(batch2$Hospital.Admission.ID))
length(unique(batch2$batch.id))
summary(batch2$batch.id)
by_id <- group_by(batch2, Hospital.Admission.ID)
check <- data.frame(summarize(by_id, num <- n()))
colnames(check) <- c("hadm.id", "num.notes")
check <- arrange(check, num.notes)
nrow(check)
head(check)
summary(check$num.notes)
#All have been annotated at least 3 times
#But do they all have at least 2 operators each?
noteop <- unique(batch2[, c("Hospital.Admission.ID", "operator")])
by_op <- noteop %>% group_by(Hospital.Admission.ID) %>% summarise(num_op <- n()) %>% data.frame()
colnames(by_op) <- c("hadm.id", "num.notes")
summary(by_op$num.notes)
#All were annotated by 3 different operators (ETM, JTW and JWW)
#Suffice to say that this batch from "JJdis22OCT15.csv" is not missing

#So the real missing batch is Miss4_id
batch4 <- alldis[alldis$Hospital.Admission.ID %in% Miss4_id,]
nrow(batch4)
length(unique(batch4$Hospital.Admission.ID))
length(unique(batch4$batch.id))
summary(batch4$batch.id)
by_id2 <- group_by(batch4, Hospital.Admission.ID)
check2 <- data.frame(summarize(by_id2, number <- n()))
colnames(check2) <- c("hadm.id", "num.notes")
check2 <- arrange(check2, num.notes)
nrow(check2)
head(check2)
summary(check2$num.notes)
# Now let's write batch4 into a suitable format for Ned's Python program
# Read in the .csv file with all the unannotated notes
summary(batch4$batch.id)
setwd("~/Documents/ICU_Frequent_Flyers-addedfiles/data/notes/dis")
fileID <- list.files(); fileID
Notesreader <- function(file_ids){
  files <- list.files()[file_ids]
  allfiles <- NULL
  for (i in 1:length(files)){
    file_i <- read.csv(files[i], sep = ",")
    colnames(file_i) <- c("subject_id", "hadm_id_MIII", "hadm_id_MIII","hadm_id_MIII", "chartdate", "category","descriptions", "text", "cohort", "randomizer")
    allfiles <- unique(rbind(allfiles, file_i))
  }
  return(allfiles)
}
#Problem files 3, 7, 8, 11, 15
miss4part1 <- Notesreader(c(2, 6, 10, 14))
miss4part2 <- Notesreader(c(7,15))
dis <- rbind(miss4part1, miss4part2)
miss4 <- dis[dis$hadm_id_MIII %in% Miss4_id, ]

#Write file to data_flyer folder
setwd("~/Documents/frequent flyer/data_flyers")
write.csv(miss4, file = "MSDis05OCT16.csv", row.names = F)
# Will send this file to Patrick and David

# Patrick and David also want me to subset "jjDis16Mar16.csv"
batch3 <- read.csv("jjDis16Mar16.csv")
nrow(batch3)
batch3_part1 <- batch3[1:25,]
batch3_part2 <- batch3[26:50,]
batch3_part3 <- batch3[51:75,]
batch3_part4 <- batch3[76:100,]
write.csv(batch3_part1, file = "jjDis16Mar16_part1.csv", row.names = F)
write.csv(batch3_part2, file = "jjDis16Mar16_part2.csv", row.names = F)
write.csv(batch3_part3, file = "jjDis16Mar16_part3.csv", row.names = F)
write.csv(batch3_part4, file = "jjDis16Mar16_part4.csv", row.names = F)
```

#Calculating the proportion of each feature that "agree" 

```{r}
#For now, we will ignore who annotated the notes (different or the same annotators in case of duplicates) and
#Define "agree" as True if all the annotations for a feature of a particular note, identified by its unique ICU.ID, are the same
#i.e. - complete agreement between all the annotated versions of a note for a feature = Agree
#We also need to exlude notes that have only been annotated once

#First, we create a function to compare the annotations of a feature 
compare_feature <- function(x, data){ #Need to use quotes when calling x.
  feature <- data[, c(x, "ICU.ID")] 
  #, "Note.ID", "Hospital.Admission.ID", "Subject.ID", "Chart.time", "cohort", "text", "Unsure", "operator")]
  feature$X <- feature[, x] #because can't use quotes with summarise
  by_icuid <- group_by(feature, ICU.ID)
  agreement <- summarise(by_icuid,
    local(num_notes <- n()),
    local(ref <- mean(X)), #column3: agree entails all 1's or all 0's
    local(pos_note <- sum(X)) 
        #column4: no. positive notes per ICU.ID for a feature
    )
  agreement <- agreement[agreement[,2] >=2,] #excl. notes annotated once
  agreement$agree <- (agreement[,3]==0 | agreement[,3]==1) 
  #return(agreement$agree) 
  return(agreement)
  #returns a T (agree) or F (disagree) vector for a feature
}

#Test function:
#Create test data frame:
test <- data.frame(ICU.ID = c(1,1,1,2,2,3,3,4), feature = c(1,0,1,0,0,1,0,1))
#Output of function as expected:
compare_feature("feature", test)
#Summary statistics of agreement:
summary(as.vector(compare_feature("feature", test)$agree))
summary(as.vector(compare_feature("feature", test)$agree)*1)
#Comment: function seems to work as it's supposed to

#Reload real dcsum:
dcsum <- read.csv("dischargeSummariesClean.csv")
#Create column names:
allfeatures <- c("None", "Obesity", "Non.Adherence", "Developmental.Delay.Retardation", "Advanced.Heart.Disease", "Advanced.Lung.Disease", "Schizophrenia.and.other.Psychiatric.Disorders", "Alcohol.Abuse", "Other.Substance.Abuse", "Chronic.Pain.Fibromyalgia", "Chronic.Neurological.Dystrophies", "Advanced.Cancer", "Depression", "Dementia", "Unsure")

#Then, we create a matrix of columns of T's and F's for all the features
AgreeMatrix <- NULL
for (i in 1:length(allfeatures)){
  logi <- as.vector(compare_feature(allfeatures[i], dcsum)$agree)
  #logi <- as.numeric(logi)
  AgreeMatrix <- cbind(AgreeMatrix, logi)
}
colnames(AgreeMatrix) <- allfeatures

#Results:
#Number of unique notes left after excluding notes that were only annotated once:
nrow(AgreeMatrix)
#Summary: tally of T's and F's for each feature:
summary(AgreeMatrix)
#Proportions of T's (Agree) and F's (Disagree):
Agreeprop <- 1*AgreeMatrix
apply(Agreeprop, 2, mean, na.rm = T)
```

#Comment:
The level of agreement is surprisingly high even after excluding notes that were annotated once...maybe I did something wrong?

This is the agreement before excluding notes that were only annotated once (which is slightly higher):
                                          None 
                                     0.7909535
                                    
                                       Obesity 
                                     0.9779951 
                                    
                                 Non.Adherence 
                                     0.9437653 
                                    
               Developmental.Delay.Retardation 
                                     0.9951100 
                                    
                        Advanced.Heart.Disease 
                                     0.8826406 
                                    
                         Advanced.Lung.Disease 
                                     0.9266504  
                                    
 Schizophrenia.and.other.Psychiatric.Disorders  
                                     0.9572127 
                                    
                                 Alcohol.Abuse 
                                     0.9572127 
                                    
                         Other.Substance.Abuse 
                                     0.9731051 
                                    
                     Chronic.Pain.Fibromyalgia 
                                     0.9254279 
                                    
              Chronic.Neurological.Dystrophies 
                                     0.8545232 
                                    
                               Advanced.Cancer 
                                     0.9633252 
                                    
                                    Depression 
                                     0.9168704 
                                    
                                      Dementia 
                                     0.9865526 
                                    
                                        Unsure 
                                     0.8312958 
                                     
                                     
#Agreement amongst positive notes
```{r}
#Identify notes with any positives - feature by feature:
identify_posnote <- function(x, data){
  y <- compare_feature(x, data)
  posnote <- y[y[, 4] >=1, ]
  #column4: number of positive notes per ICU.ID for a feature
  return(posnote)
}
posnote_num <- NULL
for(i in 1:length(allfeatures)){
  w <- nrow(identify_posnote(allfeatures[i], dcsum))
  posnote_num <- append(posnote_num, w)
}
#Proportion of agreement for each feature amongst notes with any positive features:
pos_agree_prop <- NULL
for(i in 1:length(allfeatures)){
  y <- mean(1*(identify_posnote(allfeatures[i], dcsum)$agree))
  pos_agree_prop <- append(pos_agree_prop, y)
}


#Doing the reverse for notes with all negatives (no positives) for a feature:
#Identify notes with NO positives - feature by feature:
identify_negnote <- function(x, data){
  y <- compare_feature(x, data)
  negnote <- y[y[, 4] < 1, ]
  #column4: number of positive notes per ICU.ID for a feature
  return(negnote)
}
negnote_num <- NULL
for(i in 1:length(allfeatures)){
  w <- nrow(identify_negnote(allfeatures[i], dcsum))
  negnote_num <- append(negnote_num, w)
}
#Proportion of agreement for each feature amongst notes with NO positive features, i.e. all negatives:
neg_agree_prop <- NULL
for(i in 1:length(allfeatures)){
  y <- mean(1*(identify_negnote(allfeatures[i], dcsum)$agree))
  neg_agree_prop <- append(neg_agree_prop, y)
}
#expect to be all ones by definition

#Dataframe for proportion of agreement amongst unique notes with any positive features:
posframe <- data.frame(concept = allfeatures, num_posnotes = posnote_num, prop_agreed = pos_agree_prop, prop_posnote = posnote_num/(posnote_num+negnote_num), num_negnotes = negnote_num, sum_posneg = posnote_num+negnote_num); posframe

write.table(posframe, file = "posnote_agreement.csv", row.names = F, sep = ",")

```                                    

#Doing the same with the latest discharge final file with the missing notes added
```{r}
#Load final discharge file
dcfinal <- read.csv("AllDischargeFinal24Oct16.csv")

#Then, we create a matrix of columns of T's and F's for all the features
agreematrix <- NULL
for (i in 1:length(allfeatures)){
  logi <- as.vector(compare_feature(allfeatures[i], dcfinal)$agree)
  #logi <- as.numeric(logi)
  agreematrix <- cbind(agreematrix, logi)
}
colnames(agreematrix) <- allfeatures

#Results:
#Number of unique notes left after excluding notes that were only annotated once:
nrow(agreematrix)
#Summary: tally of T's and F's for each feature:
summary(agreematrix)
#Proportions of T's (Agree) and F's (Disagree):
agreeprop <- 1*agreematrix
apply(agreeprop, 2, mean, na.rm = T)

#Identify notes with any positives - feature by feature:
identify_posnote <- function(x, data){
  y <- compare_feature(x, data)
  posnote <- y[y[, 4] >=1, ]
  #column4: number of positive notes per ICU.ID for a feature
  return(posnote)
}
posnote_Num <- NULL
for(i in 1:length(allfeatures)){
  w <- nrow(identify_posnote(allfeatures[i], dcfinal))
  posnote_Num <- append(posnote_Num, w)
}
#Proportion of agreement for each feature amongst notes with any positive features:
Pos_agree_prop <- NULL
for(i in 1:length(allfeatures)){
  y <- mean(1*(identify_posnote(allfeatures[i], dcfinal)$agree))
  Pos_agree_prop <- append(Pos_agree_prop, y)
}


#Doing the reverse for notes with all negatives (no positives) for a feature:
#Identify notes with NO positives - feature by feature:
identify_negnote <- function(x, data){
  y <- compare_feature(x, data)
  negnote <- y[y[, 4] < 1, ]
  #column4: number of positive notes per ICU.ID for a feature
  return(negnote)
}
negnote_Num <- NULL
for(i in 1:length(allfeatures)){
  w <- nrow(identify_negnote(allfeatures[i], dcfinal))
  negnote_Num <- append(negnote_Num, w)
}
#Proportion of agreement for each feature amongst notes with NO positive features, i.e. all negatives:
Neg_agree_prop <- NULL
for(i in 1:length(allfeatures)){
  y <- mean(1*(identify_negnote(allfeatures[i], dcfinal)$agree))
  Neg_agree_prop <- append(Neg_agree_prop, y)
}
#expect to be all ones by definition

#Dataframe for proportion of agreement amongst unique notes with any positive features:
Posframefinal <- data.frame(concept = allfeatures, num_posnotes = posnote_Num, prop_agreed = Pos_agree_prop, prop_posnote = posnote_Num/(posnote_Num+negnote_Num), num_negnotes = negnote_Num, sum_posneg = posnote_Num+negnote_Num); Posframefinal

write.table(Posframefinal, file = "posnote_agreementfinal.csv", row.names = F, sep = ",")

```

#Sorting and subsetting out unsures (changed to dcfinal)

```{r}
#I want to order the notes by ICU.ID then by operator before looking at the unsures and the inter & intra operator disagreements

#Create a Note ID column for the whole dataset
#So I don't lose any rows/notes even if the same operator annotated the same notes the same way
dcfinal$Note.ID <- NULL
for(i in 1:nrow(dcfinal)){
  dcfinal$Note.ID[i] <- as.numeric(rownames(dcfinal[i,]))
}

#Order by ICU.ID and then by operator 
dcfinal <- dcfinal[with(dcfinal, order(ICU.ID, operator)), ]
head(dcfinal[,c("ICU.ID", "operator", "Note.ID")])

#Subsetting out any note with "Unsure" == 1:
unsure <- dcfinal[dcfinal$Unsure == 1, ]
#Number of Unsure notes = 241
nrow(unsure)
#Number of unique notes ticked for Unsure = 203
unsureID <- unique(unsure$ICU.ID)
length(unsureID)
#Unsure by annotator
summary(unsure$operator)
#Export all unique ICU.ID for notes ticked as Unsure
write.csv(unsure, file = "unsuredcfinal.csv")


#Ned had actually sent Patrick a batch of Unsure discharge notes to adjudicate earlier:
patsure <- read.csv("PATunsureDis16Feb16Results.csv", sep = ",")
#This is only 100 of the unsure notes
str(patsure)
PTunsureID <- unique(patsure$ICU.ID)
#But only 86 unique ICU.ID's...ie there were duplicates
#Or according to Yeran, there are different note texts that share the same ICU.ID
length(unique(PTunsureID))
length(unique(patsure$Hospital.Admission.ID))
length(unique(patsure$Chart.time))
length(unique(patsure$Subject.ID))
#There's also some intra-operator variation
nrow(unique(patsure)) > length(unique(PTunsureID))


#ICU.ID of notes already adjudicated by Patrick:
adju <- PTunsureID %in% unsureID
PTinTotal <- sum(1*adju); PTinTotal
adju2 <- unsureID %in% PTunsureID
TotalinPT <- sum(1*adju2); TotalinPT
#Comment: Patrick adjudicated 64 out of the 290 unique notes in unsureID. Not sure where Ned got the other 86-64=22 notes from.

#ICU.ID of the 22 notes adjudicated by Patrick but not in our unsureID subset:
adju3 <- !(PTunsureID %in% unsureID)
adju3ID <- PTunsureID[adju3]; length(adju3ID)

#ICU.ID of unsure notes that still need adjudicating:
unadju <- !(unsureID %in% PTunsureID)
adjuID <- unsureID[unadju]; length(adjuID)

#Load all discharge notes compiled by Ned (though I suspect this set might be missing some that Ned has sent out)
#allDis <- read.csv("allDisFinal.csv") #doesn't have ICU.ID :(
#RemainUnsureDis <- allDis[allDis$ICU.ID %in% adjuID,] 
#allDis doesn't have ICU.ID so doesn't work! :(
#only has hadm_id
#let's use dcfinal instead

#Probably have to make colnames consistent for Ned's python program will still work
#Column names for a sample file from Ned:
unsuredis16feb16 <- read.csv("unsureDis16Feb16.csv")
colnames(unsuredis16feb16)

#Don't have chartdate and descriptions columns in dcfinal
#Will replace with chartdate with "Real.time" in dcfinal 
#Will create a vector of descriptions ("Summary")
RemainUnsureDis <- unique(dcfinal[dcfinal$ICU.ID %in% adjuID, c("subject.id", "Hospital.Admission.ID", "Real.time", "text")])
RemainUnsureDis$category <- rep("Discharge", nrow(RemainUnsureDis))
RemainUnsureDis$descriptions <- rep("Summary", nrow(RemainUnsureDis))

#Rename columns
colnames(RemainUnsureDis) <- c("subject_id", "hadm_id_MIII", "chartdate", "text", "category", "descriptions")
#We somehow need 2 more hadm_id columns
RemainUnsureDis$hadm_id_MIII <- RemainUnsureDis$hadm_id_MIII
RemainUnsureDis$hadm_id_MIII <- RemainUnsureDis$hadm_id_MIII
#Re-arrange the columns
RemainUnsureDis <- RemainUnsureDis[, c("subject_id", "hadm_id_MIII", "hadm_id_MIII", "hadm_id_MIII", "chartdate", "category", "descriptions", "text")]

#Need to split the remaining unsures into smaller batches for annotating:
#Saving it to unsure note folder

#setwd("~/Documents/ICU_Frequent_Flyers-addedfiles/data/notes/unsure_dis")
#write.csv(RemainUnsureDis[1:30,], file = "unsureDis25Oct16_30.csv", row.names = F)
#write.csv(RemainUnsureDis[31:60,], file = "unsureDis25Oct16_60.csv", row.names = F)
#write.csv(RemainUnsureDis[61:90,], file = "unsureDis25Oct16_90.csv", row.names = F)
#write.csv(RemainUnsureDis[91:120,], file = "unsureDis25Oct16_120.csv", row.names = F)
#write.csv(RemainUnsureDis[121:150,], file = "unsureDis25Oct16_150.csv", row.names = F)
#write.csv(RemainUnsureDis[151:180,], file = "unsureDis25Oct16_180.csv", row.names = F)
#write.csv(RemainUnsureDis[181:210,], file = "unsureDis25Oct16_210.csv", row.names = F)
#write.csv(RemainUnsureDis[211:240,], file = "unsureDis25Oct16_240.csv", row.names = F)
#write.csv(RemainUnsureDis[241:270,], file = "unsureDis25Oct16_270.csv", row.names = F)
#write.csv(RemainUnsureDis[271:311,], file = "unsureDis25Oct16_311.csv", row.names = F)
```

#let's tease out notes that disagree:

```{r}
#First create a matrix of all disagree's, along with ICU.ID identifier
DisagreeMatrix <- compare_feature("None", dcfinal)$ICU.ID
for (i in 1:length(allfeatures)){
  logi <- -1*(compare_feature(allfeatures[i], dcfinal)$agree)+1
  DisagreeMatrix <- cbind(DisagreeMatrix, logi)
}
id_col <- "ICU.ID"
colnames(DisagreeMatrix) <- append(id_col, allfeatures)
#Tease out the ICU.ID of each unique note with any disagreement
#By creating a sum_disagree column for each row
sum_disagree <- rowSums(DisagreeMatrix[, 2:16])
DisagreeMatrix <- cbind(DisagreeMatrix, sum_disagree)
nrow(DisagreeMatrix)
#Keep only notes with disagreement
DisagreeOnly <- DisagreeMatrix[DisagreeMatrix[, 17] >0,]
nrow(DisagreeOnly)

#ICU.ID of notes with any disagreement
disagreeID <- unique(DisagreeOnly[,1]); length(disagreeID)
#Number of unique ICU.ID in dcfinal
length(unique(dcfinal$ICU.ID))
#Percentage of notes that disagree is 62.89%!
paste(round(length(disagreeID)/length(unique(dcfinal$ICU.ID))*100,2), "%")

```

#Still Need to Look at Nursing Notes:

